Mathematical Gene Network Evolution

### 2017-01-29 kalman_walk.tex
edits

\documentclass[11 pt]{article}
\usepackage{amsmath, amssymb, color, xcolor}
\usepackage{graphicx}
\usepackage{fullpage}
\usepackage[hidelinks]{hyperref}

@@

$\unobservable$

@@

% is y the phenotype or H(z) the phenotype? Do we need new jargon anyway? 
%

@@

\begin{example}[Example: Oscillator.]
Motivate the system:
\begin{align*}
    A &= \left[\begin{matrix} 
        0 & 1 \\ 
       -1 & 0 
    \end{matrix}\right]
\end{align*}
as a special case of the below.

@@

>>>>

\documentclass[11 pt]{article}
\usepackage{amsmath, amssymb, color, xcolor}
\usepackage{graphicx}
\usepackage{fullpage}
\usepackage[hidelinks]{hyperref}
\usepackage{tikz}
\usetikzlibrary{arrows.meta}
\usepackage{caption, subcaption}

@@

%$\unobservable$

@@

% is y the phenotype or H(z) the phenotype? Do we need new jargon anyway? 
% 
\title{Mathematical Gene Network Evolution}
\maketitle

\begin{abstract}
Abstract.
\end{abstract}

@@

\begin{example}[Example: Oscillator.] \label{ex:osc}
Motivate the system:
\begin{align*}
    A &= \left[\begin{matrix} 
        0 & 1 \\ 
       -1 & 0 
    \end{matrix}\right]
\end{align*}

\begin{tikzpicture}
\begin{scope}[every node/.style={circle,thick,draw}]
  \node (A) at (0,0) {$x_{1}$};
    \node[dashed] (B) at (3,0) {$x_{2}$};
    \node[shape=rectangle, dashed] (U) at (1.5,-2) {$u(t)$};
\end{scope}

\begin{scope}[>={Stealth[black]},
              every node/.style={fill=white,circle},
              every edge/.style={draw=black, thick}]
    \path [->, >=Rectangle] (A) edge[bend left] node {\tiny $-1$} (B);
    \path [->] (B) edge[bend left] node {\tiny $1$} (A); 
    \path[->] (U) edge[dashed] node {\tiny $1$} (A);
    \path[->] (U) edge[dashed] node {\tiny $1$} (B);
\end{scope}
\begin{scope}[>={Stealth[black]},
              every edge/.style={draw=black, thick}]
    %\path [->] (A) edge[loop left] node {\tiny $\lambda_{1}$} (A);
    %\path [->] (B) edge[loop left] node {\tiny $\lambda_{2}$} (B);
\end{scope}

\end{tikzpicture}

as a special case of the below.

@@

\begin{example}[Hyrbid Inviability]
  Let, \begin{align*}
    A(0) =\left[\begin{matrix}
      0 & 1 \\
      -1 & 0
    \end{matrix}\right]
    \text{and }
    A(2) = \left[\begin{matrix}
      2 & -1 \\
      5 & -2
    \end{matrix}\right]
  \end{align*}
  from Example \ref{ex:osc} ($B=\begin{bmatrix} 1 \\ 1 \end{bmatrix}$ and $C=\begin{bmatrix} 1 & 0 \end{bmatrix}$). 
If hybrids are formed by exchanging genes (swapping rows), the hybrid's gene network is,
  \begin{align*}
    \text{hybrid network } := \mathcal{H}(A(0), A(2)) = \left[\begin{matrix}
      2 & -1 \\
      -1 & 0
    \end{matrix}\right] ,
  \end{align*}
  \jss{ Or should it be: \begin{align*}
    \text{hybrid network } := \text{mean}(A(0), A(2)) = \left[\begin{matrix}
      1 & 0 \\
      2 & -1
    \end{matrix}\right] ,
  \end{align*} }
  It's transfer function will be $H(z, \mathcal{H}) = \frac{z-2}{z^2 - 3z +1}$, which produces very different dynamics than its parents, $H(z, A(0)) = H(z, A(2)) = \frac{z+1}{z^2 + 1}$. The time dynamcs of molecular species $1$ in the hybrid is $y^{(\mathcal{H})} =e^{t} \cosh(\sqrt{2} t)$ \jss{ which is approx. $\approx e^{t}$}, with the concentration of species $1$ rapidly heading towards infinity. The dynamics of species $1$ in the parental networks is $y = \sin(t) + \cos(t)$, where it continuously oscillates.

  \jss{The transfer function $H(z, \mathcal{H})$, by averaging networks, is $\frac{1}{z-1}$ and the time dynamics are: $y^{(\mathcal{H})} = e^{t}$.}
\end{example}

\begin{figure}
  \centering
  \begin{subfigure}{0.5\textwidth}
    \centering
\begin{tikzpicture}
\begin{scope}[every node/.style={circle,thick,draw}]
  \node (A) at (0,0) {$x_{1}$};
    \node[dashed] (B) at (0,-4) {$x_{2}$};
    \node[shape=rectangle, dashed] (U) at (3,-2) {$u(t)$};
\end{scope}

\begin{scope}[>={Stealth[black]},
              every node/.style={fill=white,circle},
              every edge/.style={draw=black, thick}]
    %\path [->] (A) edge[bend left] node {$c$} (B);
    \path [->] (B) edge node {\tiny $\gamma$} (A); 
    \path[->] (U) edge[dashed] node {\tiny $1$} (A);
    \path[->] (U) edge[dashed] node {\tiny $1$} (B);
\end{scope}
\begin{scope}[>={Stealth[black]},
              every edge/.style={draw=black, thick}]
    \path [->] (A) edge[loop left] node {\tiny $\lambda_{1}$} (A);
    \path [->] (B) edge[loop left] node {\tiny $\lambda_{2}$} (B);
\end{scope}

\end{tikzpicture}
    \caption{Gene Regulatory Network $A(0)$.}
  \end{subfigure}%
  \begin{subfigure}{0.5\textwidth}
    \centering
\begin{tikzpicture}
\begin{scope}[every node/.style={circle,thick,draw}]
  \node (A) at (0,0) {$x_{1}$};
    \node[dashed] (B) at (0,-4) {$x_{2}$};
    \node[shape=rectangle, dashed] (U) at (3,-2) {$u(t)$};
\end{scope}

\begin{scope}[>={Stealth[black]},
              every node/.style={fill=white,circle},
              every edge/.style={draw=black, thick}]
    \path [->, sloped] (A) edge[bend left] node {\tiny $(\lambda_{2} - \lambda_{1} - \frac{\gamma}{2})$} (B);
    \path [->, sloped] (B) edge[bend left] node {\tiny $\frac{\gamma}{2}$} (A); 
    \path[->] (U) edge[dashed] node {\tiny $1$} (A);
    \path[->] (U) edge[dashed] node {\tiny $1$} (B);
\end{scope}
\begin{scope}[>={Stealth[black]},
              every edge/.style={draw=black, thick}]
    \path [->] (A) edge[loop left] node {\tiny $(\lambda_{1} + \frac{\gamma}{2})$} (A);
    \path [->] (B) edge[loop left] node {\tiny $(\lambda_{2}-\frac{\gamma}{2})$} (B);
\end{scope}

\end{tikzpicture}
    \caption{Gene Regulatory Network $A(-1)$.}
  \end{subfigure}
  \caption{Neutral rewiring of gene network $A(p)$ from Example \ref{ex:2x2}. Not only are the regulatory coefficients different between $A(0)$ and $A(-1)$, there is also a new regulatory connection.}
\end{figure}
